Notes on non-abelian generalization of Benjamin-Ono equation

### 2014-04-17 benjamin-ono/notes.tex

\end{document}

>>>>

The equation can be quantized. Mass is then changed to $\beta-\hbar$, where $\beta=\beta' \hbar$ and
$\beta-\hbar=(\beta'-1)\hbar$. (?)

The Hamiltonian is
\begin{equation}
  \label{eq:19}
  H=\frac{1}{2}\int \left(u^{2}+\frac{1}{3} u^{2} u^{H}_{x}\right) dx
\end{equation}
Now to quantize the equation we assume that function $u(x,t)$ is defined on the circle or radius
$\frac{L}{2\pi}$. Then we can expand $u$ in modes
\begin{equation}
  \label{eq:20}
  u(x)=\sum_{k} e^{i k x}a_{k}
\end{equation}
The coefficients $a_{k}$ satisfy commutation relation
\begin{equation}
  \label{eq:21}
  \{a_{m},a_{n}\}=n\delta_{nm}
\end{equation}

Equation can be written in Hamiltonian form
\begin{equation}
  \label{eq:22}
  u_{t}=\{H,u\},
\end{equation}
and
\begin{equation}
  \label{eq:23}
  H=\sum_{k>0} a_{-k} L_{k},
\end{equation}
where
\begin{equation}
  \label{eq:24}
  L_{k}=\sum_{n>0} a_{n+k} a_{n}+\beta\sum_{n>0} n a_{n+k}
\end{equation}
The functions $L_{k}$ satisfy commutation relations w.r.t. Poisson bracket of Virasoro algebra
\begin{equation}
  \label{eq:25}
  \{L_{k},L_{l}\}=(k-l)L_{k+l}+\frac{c}{12} k(k^{2}-1) \delta_{k+l,0},
\end{equation}
where
\begin{equation}
  \label{eq:26}
  c=-6\beta^{2}
\end{equation}
So this is Poisson-Lie realization of Virasoro algebra (classical Virasoro algebra).

Quantization is straightforward:
\begin{equation}
  \label{eq:27}
  [a_{m}, a_{n}]=\hbar n \delta_{nm}
\end{equation}
\begin{equation}
  \label{eq:28}
  H=\sum_{k>0}:a_{-k}L_{k}:
\end{equation}
An we obtain commutation relations of Virasoro algebra
\begin{equation}
  \label{eq:29}
   [L_{k},L_{l}]=(k-l)L_{k+l}+\frac{c}{12} k(k^{2}-1) \delta_{k+l,0},
\end{equation}
but now central charge is different
\begin{equation}
  \label{eq:30}
  c=1-6\left( \sqrt{\beta} -\frac{1}{\sqrt{\beta}}\right)^{2}
\end{equation}

Benjamin-Ono equation is also connected to random matrices. Steady state equation $u_{t}=0$ is
equivalent to loop equation for Hermitian matrices.

Quantization of solitons is somehow connected with conformal field theory. In fractional quantum
Hall waves on the boundary are described by Benjamin-Ono equation \cite{?Wiegmann}.

In conformal field theory Virasoro symmetry can be extended to Lie groups. For example, if
$u=\partial_{x} \varphi$ and $e^{i\varphi}$ is position of a particle on a ring, then $u$ is
particle displacement. Since
\begin{equation}
  \label{eq:31}
  e^{i\varphi_{1}}e^{i\varphi_{2}}=e^{i(\varphi_{1}+\varphi_{2})}
\end{equation}
we deal with $U(1)$ group.

The Benjamin-Ono equation for $\varphi$ is
\begin{equation}
  \label{eq:35}
  \partial_{t} \varphi +\frac{1}{2} (\partial_{x}\varphi)^{2}+\partial_{x}^{2}\varphi^{H}=0
\end{equation}

Now consider $\varphi\in su(2)$. Then $g=e^{i\varphi}$ is Lie group element and
\begin{equation}
  \label{eq:32}
  \hat u=-ig^{-1} \partial_{x} g\in su(2)
\end{equation}
is connection in line bundle. It is $su_{k}(2)$-current $\hat J$.
\begin{equation}
  \label{eq:33}
  J(x)=\sum \hat J_{n} e^{inx}
\end{equation}
Elements $J_{n}$ generate Kac-Moody algebra
\begin{equation}
  \label{eq:34}
  [J_{n},J_{m}]=(n-m)J_{n+m} + k n \delta_{n+m,0}
\end{equation}

The problem is to construct non-abelian extension of Benjamin-Ono equation.

If $g\leftrightarrow e^{i\varphi}$, then $\partial_{x}\varphi\leftrightarrow g^{-1}\partial_{x}g=J$.
Using (\ref{eq:35}) we can conjecture form of the generalized equation
\begin{equation}
  \label{eq:36}
  \partial_{t}J+\partial_{x}[J\cdot J]+\dots=0
\end{equation}
Such equation is supposed to have solutions with quantized mass.

Note also that Benjamin-Ono equation is a certain limit of intermediate long-wave equation (ILW).
The term $u^{H}=\fint \frac{u}{x-x'}dx'$ is the limit $L\to\infty$ of
\begin{equation}
  \label{eq:37}
  u^{H}=\fint u \coth \frac{x-x'}{L} dx'
\end{equation}
for $2L$-periodic ocean. Hilbert transform connects real and imaginary parts of complex function
$(\Re f)^{H}\to \Im f$. For complex functions the equation should be like
\begin{equation}
  \label{eq:38}
  f_{t}=f f_{z}+if_{zz}, \quad ix=z
\end{equation}
For $\varphi=\sum_{k}\frac{a_{k}}{k} e^{ikx}$ we have stress-energy tensor
\begin{equation}
  \label{eq:39}
  T=\sum_{k}L_{k}z^{-k-2}=T_{xx}+T_{yy}-2i T_{xy}=(\partial \varphi)^{2} +i\alpha\partial^{2}\varphi,
\end{equation}
and
\begin{equation}
  \label{eq:40}
  [a_{k},a_{n}]=\delta_{n+k,0},
\end{equation}
\begin{equation}
  \label{eq:41}
  L_{n}=\sum_{k}:a_{k}a_{n-k}:
\end{equation}
\begin{equation}
  \label{eq:42}
  H=\sum_{k}:a_{-k}L_{k}:
\end{equation}

Benjamin-Ono equation is equivalent to
\begin{equation}
  \label{eq:43}
  \partial_{t}u=[H,u]
\end{equation}
But we have boundary condition $a^{\dagger}_{k}=a_{-k}$. For $su(2)$-generalization we have
$a_{k}\to J_{k}, J=g \partial g$, but on the boundary we have coset $su(2)/u(1)\approx S^{2}$. And
$J=J_{x}\sigma_{x}+J_{y}\sigma_{y}+J_{z}\sigma_{z}$ should be changed to $\sigma_{z}\cdot J$. So if
we want to get Benjamin-Ono equation for spin waves we need to rewrite $H=\sum_{k}:J_{-k} L_{k}:$ on
the boundary.  
\end{document}
